Structured Versus Decorated Cospans

### 2021-06-29 cospans_new.tex
attempted to deal with all of Kenny's June 28 corrections

% Structured versus decorated cospans
% John Baez, Kenny Courser and Christina Vasilakopoulou
% in LaTeX
% 2021/6/23 - Finishing up

@@

However, we often want the system itself to have more structure than its interfaces.   This led Fong to develop a theory of `decorated' cospans \cite{Fong}.  Given a category $\A$ with finite colimits, a symmetric lax monoidal functor $F \maps (\A,+) \to (\textsf{Set},\times)$ can be used to equip the apex $m$ of a cospan in $\A$ with some extra data: an element $x \in F(m)$, which we call a \textbf{decoration}.  Thus a \define{decorated cospan} is a pair:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$m$};
\node (C) at (2,0) {$b$,};
\node (E) at (4,0) {$x \in F(m)$.};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
Fong proved that there is a symmetric monoidal category with objects
of $\A$ as its objects and equivalence classes of decorated cospans as its morphisms.  Such categories were used to describe a variety of open systems: electrical circuits, Markov processes, chemical reaction networks and dynamical systems \cite{BF,BFP,BP}.

@@

Here we present a solution to this problem.  Instead of basing the theory of decorated cospans on a symmetric lax monoidal functor $F \maps (\A, +) \to (\Set, \times)$, we use a symmetric lax monoidal pseudofunctor $F \maps (\A, +) \to (\Cat, \times)$.  In \cref{thm:decorated_cospans_1,thm:decorated_cospans_2}, we use this
data to construct a symmetric monoidal double category $F\lCsp$ in which:
\begin{itemize}
\item an object is an object of $\A$,
\item a vertical 1-morphism is a morphism of $\A$,
\item a horizontal 1-cell from $a$ to $b$ is a decorated cospan:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$m$};
\node (C) at (2,0) {$b,$};
\node (D) at (3,0) {$x \in F(m)$,};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
\item a 2-morphism is a \define{map of decorated cospans}: that is, a commutative
diagram
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0.5) {$a$};
\node (A') at (0,-0.5) {$a'$};
\node (B) at (1,0.5) {$m$};
\node (C) at (2,0.5) {$b$};
\node (C') at (2,-0.5) {$b'$};
\node (D) at (1,-0.5) {$m'$};
\node (E) at (3,0.5) {$x \in F(m)$};
\node (F) at (3,-0.5) {$x' \in F(m')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[right]{$g$} (C')
(A') edge node[above] {$i'$} (D)
(C') edge node[above] {$o'$} (D)
(B) edge node [left] {$h$} (D);
\end{tikzpicture}
\]
together with a morphism $\tau \maps F(h)(x) \to x'$ in $F(m')$.
\end{itemize}

@@

In this section we build symmetric monoidal double categories of decorated cospans, and then study the functoriality of this construction.   \cref{thm:decorated_cospans_1} explains how to construct a double category of decorated cospans from a lax monoidal pseudofunctor $F \maps (\A,+) \to (\Cat, \times)$ whenever $A$ has finite colimits.  \cref{thm:decorated_cospans_2} gives conditions under which this double category is symmetric monoidal.   These results build on earlier work of Fong \cite{Fong}, and we recommend his thesis for intuitive explanations of the fundamental ideas \cite{FongThesis}.

@@

\begin{itemize}
\item an object is an object of $\A$,
\item a vertical 1-morphism is a morphism of $\A$,
\item a horizontal 1-cell is a pair of objects $a,b \in \A$ together with an
object $c \in \C$ with $\Phi(c) = a + b$,
\item a 2-morphism is a pair of morphisms $f \maps a \to a'$, $g \maps b \to b'$ 
in $\A$ together with a morphism $h \maps c \to c'$ with $\Phi(h) = f + g$.
\end{itemize}
If $\Phi$ is a Beck--Chevalley symmetric monoidal bifibration then $\Fr(\Phi)$ becomes a symmetric
monoidal double category.
\end{lem}

\begin{proof} This is \cite[Theorem 14.9]{Shulman2008}; Shulman proves the dual version in more detail in his Theorem 14.4.  
\end{proof}

@@

We described $\A/U$ and its monoidal structure in the proof of \cref{thm:decorated_cospans_1}, 
and of course the case of $\A'/U'$ is analogous.   We define $G \maps \A/U \to \A'/U'$ as follows.  On objects $G$ is given by
\[  G(a \xrightarrow{i} b, s) \; = \; (H(a) \xrightarrow{H(i)} H(b), \underline{s}), \]
where to define $\underline{x}$ we treat $s \in F(b)$ as a functor 
$s \maps 1 \to F(b)$ and let $\underline{s} \in F'(H(b))$ be the composite functor
\[  1 \xrightarrow{\phi_0} E(1) \xrightarrow{E(x)} E(F(b)) \xrightarrow{\;\theta_b\;} F'(H(b)) .\]
On morphisms $(f,g,h) \maps (a \xrightarrow{i} b, s) \; \to \; 
(a' \xrightarrow{i'} b', s')$, $G$ is given by
\[   G(f,g,h) = (H(f),H(g),\underline{h}) \]
where to define $\underline{h}$ we treat $h \maps F(g)(s) \to s'$ as a natural transformation
\[
\begin{tikzpicture}[scale=1.5]
\node (G) at (-3.5,-0.5) {$1$};
\node (H) at (-2,0) {$F(b)$};
\node (I) at (-2,-1) {$F(b')$};
\node (J) at (-2.5,-0.5) {$\scriptstyle\Downarrow\tau$};
\path[->,font=\scriptsize,>=angle 90]
(G) edge node [above] {$s$} (H)
(G) edge node [below] {$s'$} (I)
(H) edge node [right] {$F(g)$} (I);
\end{tikzpicture}
\]
and let $\underline{h} \maps F'(H(g))(\underline{s}) \to \underline{s}'$ be this
natural transformation:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (-0.25,-0.5) {$1$};
\node (L) at (0.5,-0.5) {$E(1)$};
\node (B) at (2,0) {$E(F(b))$};
\node (D) at (2,-1) {$E(F(b'))$};
\node (C) at (1.5,-0.5) {$\scriptstyle\Downarrow E(\tau)$};
\node (E) at (4,0) {$F^\prime(H(b))$};
\node (F) at (4,-1) {$F^\prime(H(b'))$.};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node [above] {$\phi_0$} (L)
(B) edge node [above] {$\theta_b$} (E)
(D) edge node [above] {$\theta_{b'}$} (F)
(E) edge node [right] {$F'(H(h))$} (F)
(L) edge node[above]{$E(s)$} (B)
(L) edge node[below]{$E(s')$} (D)
(B) edge node[right]{$E(F(h))$} (D);
\end{tikzpicture}
\]
One can check that with these definitions $G$ is a functor.

@@

The comma category $L/\X$ has finite colimits because $A$ and $X$ have finite colimits and $L$ 
preserves them.   It thus becomes symmetric monoidal with the chosen finite coproducts providing the monoidal structure.   The tensor product of two objects $(a,L(a) \xrightarrow{i} x)$ and $(a,L(a) \xrightarrow{i'} x')$ is given by 
\[  (a,L(a) \xrightarrow{i} x) + (a',L(a') \xrightarrow{i'} x')=(a+a',L(a+a')  \xrightarrow{\sim} L(a)+L(a') \xrightarrow{i+i'} x+x'). \]
The monoidal unit is ${(0_\A,0_\X,{L(0_\A) \xrightarrow{!} 0_\X})}$.  It is clear that $\Psi \maps (L/\X, +) \to (\A,+)$ is a strict monoidal functor.

@@

As a left adjoint, $L$ preserves all colimits that exist between the categories $\A$ and $\inta F$, which have finite colimits.   Thus, we can construct the symmetric monoidal double category of structured cospans $_L\lCsp(\inta F)$ as in \cref{thm:structured_cospans}.  To show that this is isomorphic to the symmetric monoidal double category of decorated cospans $F\lCsp$ given by  \cref{thm:decorated_cospans_2}, we use a result of Shulman \cite[Theorem~14.11]{Shulman2010}.    Namely, the $\mathbb{F}\mathbf{r}$ construction of \cref{lem:shulman} extends to a 2-functor 
\[ \mathbb{F}\mathbf{r} \maps \mathbf{FMF} \to \mathbf{FDbl} \]
where $\mathbf{FMF}$ is the 2-category consisting of
\begin{itemize}
\item Beck--Chevalley monoidal bifibrations,
\item strong monoidal morphisms of bifibrations, and
\item monoidal transformations of bifibrations
\end{itemize}
while $\mathbf{FDbl}$ is the 2-category of 
\begin{itemize}
\item fibrant double categories, 
\item double functors, and 
\item double transformations. 
\end{itemize}
Given two monoidal bifibrations $\Phi \maps \A \to \mathsf{B}$ and $\Phi' \maps \A' \to \mathsf{B}'$, a `strong monoidal morphism of bifibrations' from $\Phi$ to $\Phi'$ consists of a pair of strong monoidal functors $F_1$ and $F_2$ making the following square commutes
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$\A$};
\node (B) at (1,0) {$\A'$};
\node (C) at (0,-1) {$\mathsf{B}$};
\node (D) at (1,-1) {$\mathsf{B}'$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$F_1$} (B)
(A) edge node[left]{$\Phi$} (C)
(B) edge node[right]{$\Phi'$} (D)
(C) edge node[below]{$F_2$} (D);
\end{tikzpicture}
\]
not just as functors but as strong monoidal functors.   In our case the two monoidal bifibrations are the functors $\Phi$ and $\Psi$ of Theorems \ref{thm:decorated_cospans_2} and \ref{thm:structured_cospans}, respectively.  These share $\A$ as a common base, so we only need a single strong monoidal functor $F \maps \A/U \to L/\X$ making this diagram of strong monoidal functors commute:
\[
\begin{tikzpicture}[scale=1.0]
\node (A) at (0,0) {$\A/U$};
\node (B) at (2,0) {$L/ \X$};
\node (C) at (1,-1.5) {$\A$.};
%\node (D) at (2,-1) {$\A$};
%\node (E) at (1,0) {$\Downarrow \alpha_1$};
%\node (F) at (1,-1) {$\Downarrow \alpha_2$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$F$} (B)
%(B) edge [bend left] node[above]{$$} (A)
(A) edge node[left]{$\Phi$} (C)
(B) edge node[right]{$\Psi$} (C);
%(C) edge [bend left] node[above]{$G_1$} (C)
%(C) edge [bend right] node[above]{$G_2$} (C);
\end{tikzpicture}
\]
There is an isomorphism $F \maps \A/U \to L/\X$ arising from the fact that $U$ is right adjoint to $L$.   It is clear that $F$ is strong monoidal via componentwise binary coproducts which are preserved by both $L$ and $U$, and that $\Psi F=\Phi$ as strong monoidal functors.  As $F$ is an isomorphism, we obtain an isomorphism 
\[  \mathbb{F}\mathbf{r}(F) \maps \mathbb{F}\mathbf{r}(\Phi) \to \mathbb{F}\mathbf{r}(\Psi)\] between the two monoidal double categories $\mathbb{F}\mathbf{r}(\Phi)=F\lCsp$ and $\mathbb{F}\mathbf{r}(\Psi)= {}_L \lCsp(\inta F)$ of Theorems \ref{thm:decorated_cospans_2} and \ref{thm:structured_cospans}, respectively.   One can check by hand that this isomorphism 
is symmetric monoidal.
\end{proof}

@@

\node (K) at (8,-0.5) {$I_a \in F(b)$};

@@

Thinking about systems and processes categorically dates back to early works by Lawvere \cite{Lawvere}, Bunge--Fiore \cite{BungeFiore}, Joyal--Nielsen--Winskel \cite{JNW}, Katis--Sabadini--Walters \cite{KSW} and others.   Spivak and others have used wiring diagrams and sheaves to capture compositional features of dynamical systems, \cite{BFV,SSV,VSL}.  Another approach uses signal flow diagrams and other string diagrams \cite{BE,BSZ,FRS} to understand systems behaviorally, following ideas of Willems \cite{Willems}.

@@

One of the simplest kinds of network is a graph.  For us a \define{graph} will be a pair of functions $s,t\maps E \to N$ where $E$ and $N$ are finite sets.   We call elements of $E$ \define{edges} and elements of $N$ \define{nodes}.  There is a category $\Graph$ where the objects are graphs and a morphism from the graph $s,t\maps E \to N$ to the graph $s',t' \maps E' \to N'$ is a pair of functions $f \maps E \to E', g \maps N \to N'$ such that these diagrams commute:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$E$};
\node (A') at (0,-1) {$E'$};
\node (B) at (1,0) {$N$};
\node (B') at (1,-1) {$N'$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$s$} (B)
(A') edge node[above]{$s'$} (B')
(A) edge node[left]{$f$} (A')
(B) edge node[right]{$g$} (B');

\node (C) at (2,0) {$E$};
\node (C') at (2,-1) {$E'$};
\node (D) at (3,0) {$N$};
\node (D') at (3,-1) {$N'$.};
\path[->,font=\scriptsize,>=angle 90]
(C) edge node[above]{$t$} (D)
(C') edge node[above]{$t'$} (D')
(C) edge node[left]{$f$} (C')
(D) edge node[right]{$g$} (D');
\end{tikzpicture}
\]

We can easily build a double category with `open graphs' as horizontal 1-cells using the machinery of structured cospans, see \cite[Section 5]{BC}.  Let $L \maps \Fin\Set \to \Graph$ be the functor that assigns to a finite set $N$ the \define{discrete graph} on $N$: the graph with no edges and $N$ as its set of vertices. Both $\Fin\Set$ and $\Graph$ have finite colimits, and the functor $L \maps \Fin\Set \to \Graph$ is left adjoint to the forgetful functor $R \maps \Graph \to \Fin\Set$ that assigns to a graph $G$ its underlying set of vertices $R(G)$. Thus, using structured cospans and appealing to \cref{thm:structured_cospans}, we get a symmetric monoidal double category $_L \lCsp(\Graph)$ in which:
\begin{itemize}
\item objects are finite sets,
\item a vertical 1-morphism from $X$ to $Y$ is a function $f \maps X \to Y$,
\item a horizontal 1-cell from $X$ to $Y$ is an \define{open graph} from $X$ to $Y$, meaning a cospan in $\Graph$ of this form:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(X)$};
\node (B) at (1,0) {$G$};
\node (C) at (2,0) {$L(Y)$,};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
\item a 2-morphism is a commuting diagram in $\Graph$ of this form:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(X_1)$};
\node (B) at (1,0) {$G_1$};
\node (C) at (2,0) {$L(Y_1)$};
\node (A') at (0,-1) {$L(X_2)$};
\node (B') at (1,-1) {$G_2$};
\node (C') at (2,-1) {$L(Y_2)$.};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$I_1$} (B)
(C) edge node[above]{$O_1$} (B)
(A') edge node[above]{$I_2$} (B')
(C') edge node[above]{$O_2$} (B')
(A) edge node [left]{$L(f)$} (A')
(B) edge node [left]{$\alpha$} (B')
(C) edge node [left]{$L(g)$} (C');
\end{tikzpicture}
\]
\end{itemize}
Here is an example of an open graph:
\[
\scalebox{0.8}{
\begin{tikzpicture}
	\begin{pgfonlayer}{nodelayer}
		\node [contact] (n1) at (-2,0) {$\bullet$};
		\node [style = none] at (-2.1,0.3) {$n_1$};
		\node [contact] (n2) at (0,1) {$\bullet$};
		\node [style = none] at (0,1.3) {$n_2$};
		\node [contact] (n3) at (0,-1) {$\bullet$};
		\node [style = none] at (0,-1.3) {$n_3$};
		\node [contact] (n4) at (2,0) {$\bullet$};
		\node [style = none] at (2.1,0.3) {$n_4$};

@@

Taking $D$ as above, it is easy to see that in the category $\inta D$
\begin{itemize}
\item an object is a pair $(S,v)$ where $S$ is a finite set and $v$ is an algebraic vector field $v \maps \R^S \to \R^S$,
\item a morphism from $(S,v)$ to $(S',v')$ is a function $f \maps S \to S'$ such that $v' = f_* \circ v \circ f^*$.
\end{itemize}
with the usual composition of functions.   The forgetful functor $U \maps \inta D \to \Fin\Set$ acts as follows:
\begin{itemize}
\item on objects, $D(S,v) = S$,
\item on morphisms, $D(f) = f$.
\end{itemize}

>>>>

% Structured versus decorated cospans
% John Baez, Kenny Courser and Christina Vasilakopoulou
% in LaTeX
% 2021/6/28 - Finishing up

@@

However, we often want the system itself to have more structure than its interfaces.   This led Fong to develop a theory of `decorated' cospans \cite{Fong}.  Given a category $\A$ with finite colimits, a symmetric lax monoidal functor $F \maps (\A,+) \to (\textsf{Set},\times)$ can be used to equip the apex $m$ of a cospan in $\A$ with some extra data: an element $s \in F(m)$, which we call a \textbf{decoration}.  Thus a \define{decorated cospan} is a pair:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$m$};
\node (C) at (2,0) {$b$,};
\node (E) at (4,0) {$s \in F(m)$.};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
Fong proved that there is a symmetric monoidal category with objects
of $\A$ as its objects and equivalence classes of decorated cospans as its morphisms.  Such categories were used to describe a variety of open systems: electrical circuits, Markov processes, chemical reaction networks and dynamical systems \cite{BF,BFP,BP}.

@@

Here we present a solution to this problem.  Instead of basing the theory of decorated cospans on a symmetric lax monoidal functor $F \maps (\A, +) \to (\Set, \times)$, we use a symmetric lax monoidal pseudofunctor $F \maps (\A, +) \to (\Cat, \times)$.  In \cref{thm:decorated_cospans_1,thm:decorated_cospans_2}, we use this
data to construct a symmetric monoidal double category $F\lCsp$ in which:
\begin{itemize}
\item an object is an object of $\A$,
\item a vertical 1-morphism is a morphism of $\A$,
\item a horizontal 1-cell from $a$ to $b$ is a decorated cospan:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$m$};
\node (C) at (2,0) {$b,$};
\node (D) at (3,0) {$s \in F(m)$,};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
\item a 2-morphism is a \define{map of decorated cospans}: that is, a commutative
diagram
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0.5) {$a$};
\node (A') at (0,-0.5) {$a'$};
\node (B) at (1,0.5) {$m$};
\node (C) at (2,0.5) {$b$};
\node (C') at (2,-0.5) {$b'$};
\node (D) at (1,-0.5) {$m'$};
\node (E) at (3,0.5) {$s \in F(m)$};
\node (F) at (3,-0.5) {$s' \in F(m')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[right]{$g$} (C')
(A') edge node[above] {$i'$} (D)
(C') edge node[above] {$o'$} (D)
(B) edge node [left] {$h$} (D);
\end{tikzpicture}
\]
together with a morphism $\tau \maps F(h)(s) \to s'$ in $F(m')$.
\end{itemize}

@@

In this section we build symmetric monoidal double categories of decorated cospans, and then study the functoriality of this construction.   \cref{thm:decorated_cospans_1} explains how to construct a double category of decorated cospans from a lax monoidal pseudofunctor $F \maps (\A,+) \to (\Cat, \times)$ whenever $\A$ has finite colimits.  \cref{thm:decorated_cospans_2} gives conditions under which this double category is symmetric monoidal.   These results build on earlier work of Fong \cite{Fong}, and we recommend his thesis for intuitive explanations of the fundamental ideas \cite{FongThesis}.

@@

\begin{itemize}
\item an object is an object of $\A$,
\item a vertical 1-morphism is a morphism of $\A$,
\item a horizontal 1-cell is a pair of objects $a,b \in \A$ together with an
object $c \in \C$ with $\Phi(c) = a + b$,
\item a 2-morphism is a pair of morphisms $f \maps a \to a'$, $g \maps b \to b'$ 
in $\A$ together with a morphism $h \maps c \to c'$ in $\C$ with $\Phi(h) = f + g$.
\end{itemize}
If $\Phi$ is a Beck--Chevalley symmetric monoidal bifibration then $\Fr(\Phi)$ becomes a symmetric
monoidal double category.
\end{lem}

\begin{proof} This is \cite[Theorem 14.9]{Shulman2008}; Shulman proves a dual version in more detail in his Theorem 14.4.  
\end{proof}

@@

We described $\A/U$ and its monoidal structure in the proof of \cref{thm:decorated_cospans_1}, 
and of course the case of $\A'/U'$ is analogous.   We define $G \maps \A/U \to \A'/U'$ as follows.  On objects $G$ is given by
\[  G(a \xrightarrow{i} b, s) \; = \; (H(a) \xrightarrow{H(i)} H(b), \underline{s}), \]
where to define $\underline{s}$ we treat $s \in F(b)$ as a functor 
$s \maps 1 \to F(b)$ and let $\underline{s} \in F'(H(b))$ be the composite functor
\[  1 \xrightarrow{\phi_0} E(1) \xrightarrow{E(s)} E(F(b)) \xrightarrow{\;\theta_b\;} F'(H(b)) .\]
On morphisms $(f,g,h) \maps (a \xrightarrow{i} b, s) \; \to \; 
(a' \xrightarrow{i'} b', s')$, $G$ is given by
\[   G(f,g,h) = (H(f),H(g),\underline{h}) \]
where to define $\underline{h}$ we treat $h \maps F(g)(s) \to s'$ as a natural transformation
\[
\begin{tikzpicture}[scale=1.5]
\node (G) at (-3.5,-0.5) {$1$};
\node (H) at (-2,0) {$F(b)$};
\node (I) at (-2,-1) {$F(b')$};
\node (J) at (-2.5,-0.5) {$\scriptstyle\Downarrow h$};
\path[->,font=\scriptsize,>=angle 90]
(G) edge node [above] {$s$} (H)
(G) edge node [below] {$s'$} (I)
(H) edge node [right] {$F(g)$} (I);
\end{tikzpicture}
\]
and let $\underline{h} \maps F'(H(g))(\underline{s}) \to \underline{s}'$ be this
natural transformation:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (-0.25,-0.5) {$1$};
\node (L) at (0.5,-0.5) {$E(1)$};
\node (B) at (2,0) {$E(F(b))$};
\node (D) at (2,-1) {$E(F(b'))$};
\node (C) at (1.5,-0.5) {$\scriptstyle\Downarrow E(h)$};
\node (E) at (4,0) {$F^\prime(H(b))$};
\node (F) at (4,-1) {$F^\prime(H(b'))$.};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node [above] {$\phi_0$} (L)
(B) edge node [above] {$\theta_b$} (E)
(D) edge node [above] {$\theta_{b'}$} (F)
(E) edge node [right] {$F'(H(g))$} (F)
(L) edge node[above]{$E(s)$} (B)
(L) edge node[below]{$E(s')$} (D)
(B) edge node[right]{$E(F(g))$} (D);
\end{tikzpicture}
\]
One can check that with these definitions $G$ is a functor.

@@

The comma category $L/\X$ has finite colimits because $\A$ and $\X$ have finite colimits and $L$ 
preserves them.   It thus becomes symmetric monoidal with the chosen finite coproducts providing the monoidal structure.   The tensor product of two objects $(a,L(a) \xrightarrow{i} x)$ and $(a,L(a) \xrightarrow{i'} x')$ is given by 
\[  (a,L(a) \xrightarrow{i} x) + (a',L(a') \xrightarrow{i'} x')=(a+a',L(a+a')  \xrightarrow{\sim} L(a)+L(a') \xrightarrow{i+i'} x+x'). \]
The monoidal unit is $(0_\A, L(0_\A) \xrightarrow{!} 0_\X)$.  It is clear that $\Psi \maps (L/\X, +) \to (\A,+)$ is a strict monoidal functor.

@@

As a left adjoint, $L$ preserves all colimits that exist between the categories $\A$ and $\inta F$, which have finite colimits.   Thus, we can construct the symmetric monoidal double category of structured cospans $_L\lCsp(\inta F)$ as in \cref{thm:structured_cospans}.  To show that this is isomorphic to the symmetric monoidal double category of decorated cospans $F\lCsp$ given by  \cref{thm:decorated_cospans_2}, we use a result of Shulman \cite[Theorem~14.11]{Shulman2010}.    Namely, the $\mathbb{F}\mathbf{r}$ construction of \cref{lem:shulman} extends to a 2-functor 
\[ \mathbb{F}\mathbf{r} \maps \mathbf{BCMF} \to \mathbf{FDbl} \]
where $\mathbf{BCMF}$ is the 2-category consisting of
\begin{itemize}
\item Beck--Chevalley monoidal bifibrations,
\item strong monoidal morphisms of bifibrations, and
\item monoidal transformations of bifibrations
\end{itemize}
while $\mathbf{FDbl}$ is the 2-category of 
\begin{itemize}
\item fibrant double categories, 
\item double functors, and 
\item double transformations. 
\end{itemize}
Given two monoidal bifibrations $\Phi \maps \A \to \mathsf{B}$ and $\Phi' \maps \A' \to \mathsf{B}'$, a `strong monoidal morphism of bifibrations' from $\Phi$ to $\Phi'$ consists of a pair of strong monoidal functors $F_1$ and $F_2$ making the following square commute
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$\A$};
\node (B) at (1,0) {$\A'$};
\node (C) at (0,-1) {$\mathsf{B}$};
\node (D) at (1,-1) {$\mathsf{B}'$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$F_1$} (B)
(A) edge node[left]{$\Phi$} (C)
(B) edge node[right]{$\Phi'$} (D)
(C) edge node[below]{$F_2$} (D);
\end{tikzpicture}
\]
not just as functors but as strong monoidal functors.   In our case the two monoidal bifibrations are the functors $\Phi$ and $\Psi$ of Theorems \ref{thm:decorated_cospans_2} and \ref{thm:structured_cospans}, respectively.  These share $\A$ as a common base, so we only need a single strong monoidal functor $F \maps \A/U \to L/\X$ making this diagram of strong monoidal functors commute:
\[
\begin{tikzpicture}[scale=1.0]
\node (A) at (0,0) {$\A/U$};
\node (B) at (2,0) {$L/ \X$};
\node (C) at (1,-1.5) {$\A$.};
%\node (D) at (2,-1) {$\A$};
%\node (E) at (1,0) {$\Downarrow \alpha_1$};
%\node (F) at (1,-1) {$\Downarrow \alpha_2$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$F$} (B)
%(B) edge [bend left] node[above]{$$} (A)
(A) edge node[left]{$\Phi$} (C)
(B) edge node[right]{$\Psi$} (C);
%(C) edge [bend left] node[above]{$G_1$} (C)
%(C) edge [bend right] node[above]{$G_2$} (C);
\end{tikzpicture}
\]
There is an isomorphism $F \maps \A/U \to L/\X$ arising from the fact that $U$ is right adjoint to $L$.   It is clear that $F$ is strong monoidal via componentwise binary coproducts which are preserved by both $L$ and $U$, and that $\Psi F=\Phi$ as strong monoidal functors.  As $F$ is an isomorphism, we obtain an isomorphism 
\[  \mathbb{F}\mathbf{r}(F) \maps \mathbb{F}\mathbf{r}(\Phi) \to \mathbb{F}\mathbf{r}(\Psi)\] between the two monoidal double categories $\mathbb{F}\mathbf{r}(\Phi)=F\lCsp$ and $\mathbb{F}\mathbf{r}(\Psi)= {}_L \lCsp(\inta F)$ of Theorems \ref{thm:decorated_cospans_2} and \ref{thm:structured_cospans}, respectively.   One can check by hand that this isomorphism 
is symmetric monoidal.
\end{proof}

@@

\node (K) at (8,-0.5) {$I_b \in F(b)$};

@@

Thinking about systems and processes categorically dates back to early works by Lawvere \cite{Lawvere}, Bunge--Fiore \cite{BungeFiore}, Joyal--Nielsen--Winskel \cite{JNW}, Katis--Sabadini--Walters \cite{KSW} and others.   Spivak and others have used wiring diagrams and sheaves to capture compositional features of dynamical systems \cite{BFV,SSV,VSL}.  Another approach uses signal flow diagrams and other string diagrams \cite{BE,BSZ,FRS} to understand systems behaviorally, following ideas of Willems \cite{Willems}.

@@

One of the simplest kinds of network is a graph.  For us a \define{graph} will be a pair of functions $s,t\maps E \to N$ where $E$ and $N$ are finite sets.   We call elements of $E$ \define{edges} and elements of $N$ \define{nodes}.  There is a category $\Graph$ where the objects are graphs and a morphism from the graph $s,t\maps E \to N$ to the graph $s',t' \maps E' \to N'$ is a pair of functions $f \maps E \to E', g \maps N \to N'$ such that these diagrams commute:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$E$};
\node (A') at (0,-1) {$E'$};
\node (B) at (1,0) {$N$};
\node (B') at (1,-1) {$N'$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$s$} (B)
(A') edge node[below]{$s'$} (B')
(A) edge node[left]{$f$} (A')
(B) edge node[right]{$g$} (B');

\node (C) at (2,0) {$E$};
\node (C') at (2,-1) {$E'$};
\node (D) at (3,0) {$N$};
\node (D') at (3,-1) {$N'$.};
\path[->,font=\scriptsize,>=angle 90]
(C) edge node[above]{$t$} (D)
(C') edge node[below]{$t'$} (D')
(C) edge node[left]{$f$} (C')
(D) edge node[right]{$g$} (D');
\end{tikzpicture}
\]

We can easily build a double category with `open graphs' as horizontal 1-cells using the machinery of structured cospans \cite[Section 5]{BC}.  Let $L \maps \Fin\Set \to \Graph$ be the functor that assigns to a finite set $N$ the \define{discrete graph} on $N$: the graph with no edges and $N$ as its set of vertices. Both $\Fin\Set$ and $\Graph$ have finite colimits, and the functor $L \maps \Fin\Set \to \Graph$ is left adjoint to the forgetful functor $R \maps \Graph \to \Fin\Set$ that assigns to a graph $G$ its underlying set of vertices $R(G)$. Thus, using structured cospans and appealing to \cref{thm:structured_cospans}, we get a symmetric monoidal double category $_L \lCsp(\Graph)$ in which:
\begin{itemize}
\item objects are finite sets,
\item a vertical 1-morphism from $X$ to $Y$ is a function $f \maps X \to Y$,
\item a horizontal 1-cell from $X$ to $Y$ is an \define{open graph} from $X$ to $Y$, meaning a cospan in $\Graph$ of this form:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(X)$};
\node (B) at (1,0) {$G$};
\node (C) at (2,0) {$L(Y)$,};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
\item a 2-morphism is a commuting diagram in $\Graph$ of this form:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$L(X)$};
\node (B) at (1,0) {$G$};
\node (C) at (2,0) {$L(Y)$};
\node (A') at (0,-1) {$L(X')$};
\node (B') at (1,-1) {$G'$};
\node (C') at (2,-1) {$L(Y')$.};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A') edge node[below]{$i'$} (B')
(C') edge node[below]{$o'$} (B')
(A) edge node [left]{$L(f)$} (A')
(B) edge node [left]{$\alpha$} (B')
(C) edge node [right]{$L(g)$} (C');
\end{tikzpicture}
\]
\end{itemize}
Here is an example of an open graph:
\[
\scalebox{0.8}{
\begin{tikzpicture}
	\begin{pgfonlayer}{nodelayer}
		\node [contact] (n1) at (-2,0) {$\bullet$};
		\node [style = none] at (-2.1,0.3) {$n_1$};
		\node [contact] (n2) at (0,1) {$\bullet$};
		\node [style = none] at (0,1.3) {$n_2$};
		\node [contact] (n3) at (0,-1) {$\bullet$};
		\node [style = none] at (0,-1.3) {$n_3$};
		\node [contact] (n4) at (2,0) {$\bullet$};
		\node [style = none] at (2.1,0.3) {$n_4$};

@@

Taking $D$ as above, it is easy to see that in the category $\inta D$
\begin{itemize}
\item an object is a pair $(S,v)$ where $S$ is a finite set and $v$ is an algebraic vector field $v \maps \R^S \to \R^S$,
\item a morphism from $(S,v)$ to $(S',v')$ is a function $f \maps S \to S'$ such that $v' = f_* \circ v \circ f^*$
\end{itemize}
with the usual composition of functions.   The forgetful functor $U \maps \inta D \to \Fin\Set$ acts as follows:
\begin{itemize}
\item on objects, $D(S,v) = S$,
\item on morphisms, $D(f) = f$.
\end{itemize}
